Main Analyses of the Combined Dataset

### 2021-04-29 manuscript/manuscript.tex
new data fixes, bug fixes, etc for revisions

One limitation of this study is the inability to quantify the impact of spatial
population genetic processes on the relationship between diversity and census
population sizes across taxa. The genomic diversity estimates collated in this
study unfortunately lack details about the sampling process and spatial data,
which can have a profound impact on population genomic summary statistics
\parencite{Battey2020-lc}. These issues could systematically bias species-wide
diversity estimates; for example, if diversity estimates from a cosmopolitan
species were primarily from a single subpopulation, diversity would be an
underestimate relative to the entire population. However, biased spatial
sampling alone seems incapable of explaining the $\pi$-$N_c$ divergence in
high-$N_c$ taxa. In the extreme scenario in which only one subpopulation was
sampled, $F_\text{ST}$ would need to be close to one for population subdivision
alone to sufficiently reduce the total population heterozygosity to explain the
orders-of-magnitude shortfall between predicted and observed diversity levels.
Yet, across-taxa surveys indicate that $F_\text{ST}$ is almost never this high
within species \parencite{Roux2016-lm}.  Still, future work could quantify the
extent to which spatial processes contribute to Lewontin's Paradox. For
example, high-$N_c$ taxa usually experience range expansions, likely with
repeated founder effects and local extinction/recolonization dynamics that
doubtlessly depress diversity. In particular, with the appropriate data, one
could estimate the empirical relationship between dispersal distance, range
size, and coalescent effective population size across taxa.

@@

I would like to thank Andy Kern and Peter Ralph for helpful discussions and
supporting me during this work, and Graham Coop for inspiration and helpful
feedback during socially distanced nature walks at Yolo Basin. I thank Jessica
Stapley for kindly providing the recombination map length data, and Yaniv
Brandvain, Amy Collins, Doc Edge, Tyler Kent, Chuck Langley, Matt Osmond, Sally
Otto, Jeff Ross-Ibarra, Aaron Stern, Anastasia Teterina, Michael Turelli, and
my Kern-Ralph labmates for helpful discussions. Sarah Friedman, Katherine Corn,
and Josef Uyeda provided very useful advice about phylogenetic comparative
methods; yet I take full responsibility for any shortcomings of my analysis.  I
would like to also thank UO librarian Dean Walton for helping me track down
some rather difficult to find older papers. This work was supported by an NIH
Grant (1R01GM117241) awarded to Andrew Kern.

@@

\section*{Supplementary Table and Figures}

>>>>

One limitation of this study is the inability to quantify the impact of spatial
population genetic processes on the relationship between diversity and census
population sizes across taxa. The genomic diversity estimates collated in this
study unfortunately lack details about the sampling process and spatial data,
which can have a profound impact on population genomic summary statistics
\parencite{Battey2020-lc}. These issues could systematically bias species-wide
diversity estimates; for example, if diversity estimates from a cosmopolitan
species were primarily from a single subpopulation, diversity would be an
underestimate relative to the entire population. However, biased spatial
sampling alone seems incapable of explaining the $\pi$-$N_c$ divergence in
high-$N_c$ taxa. In the extreme scenario in which only one subpopulation was
sampled, $F_\text{ST}$ would need to be close to one for population subdivision
alone to sufficiently reduce the total population heterozygosity to explain the
orders-of-magnitude shortfall between predicted and observed diversity levels.
This is because the equation for $F_{ST}$ can be rearranged such that $H_S =
(1-F_{ST})H_T$, where $H_S$ and $H_S$ are the subpopulation and total
population heterozygosities; if $H_T = 4N_c\mu$, then only $F_{ST} \approx 1$
can reduce $H_S$ several orders of magnitude. Yet, across-taxa surveys indicate
that $F_\text{ST}$ is almost never this high within species
\parencite{Roux2016-lm}.  Still, future work could quantify the extent to which
spatial processes contribute to Lewontin's Paradox. For example, high-$N_c$
taxa usually experience range expansions, likely with repeated founder effects
and local extinction/recolonization dynamics that doubtlessly depress
diversity. In particular, with the appropriate data, one could estimate the
empirical relationship between dispersal distance, range size, and coalescent
effective population size across taxa.

@@

I would like to thank Andy Kern and Peter Ralph for helpful discussions and
supporting me during this work, and Graham Coop for inspiration and helpful
feedback during socially distanced nature walks at Yolo Basin. I thank Jessica
Stapley for kindly providing the recombination map length data, and Yaniv
Brandvain, Amy Collins, Doc Edge, Tyler Kent, Chuck Langley, Matt Osmond, Sally
Otto, Jeff Ross-Ibarra, Aaron Stern, Anastasia Teterina, Michael Turelli,
Margot Wood, and my Kern-Ralph labmates for helpful discussions. Sarah
Friedman, Katherine Corn, and Josef Uyeda provided very useful advice about
phylogenetic comparative methods; yet I take full responsibility for any
shortcomings of my analysis.  I would like to also thank UO librarian Dean
Walton for helping me track down some rather difficult to find older papers.
This work was supported by an NIH Grant (1R01GM117241) awarded to Andrew Kern.

@@

\section*{Supplementary Information}

\section{Population Size Validation}

### 2021-04-29 notebooks/main_analysis.Rmd
new data fixes, bug fixes, etc for revisions

@@

@@

# https://bionumbers.hms.harvard.edu/bionumber.aspx?&id=107256 
# 18-23% of human body is carbon; seems reasonable this holds
# for other animals

@@

# in Gt
prop_carbon <- signif(sqrt(0.18 * 0.23), 1)  # ~ 0.2
# 1Gt = 1e12 kg
xd$carbon_mass <- (prop_carbon*(10^xd$pred_log10_body_mass) / 1e3 * 10^xd$log10_popsize) / 1e12

mass_d <- xd %>% 
  # filter so we don't include NA species in nspecies
  filter(is.finite(carbon_mass), is.finite(log10_popsize)) %>%
  group_by(phylum) %>% 
  summarize(total_carbon_mass=sum(carbon_mass, na.rm=TRUE),
            total_log10_popsize=log10(sum(10^log10_popsize, na.rm=TRUE)),
            nspecies = n()) %>%
  mutate(total = sum(total_carbon_mass), prop_carbon_mass = total_carbon_mass / total)

# chordates are incredibly over-represented in my data set by biomass
baron_d <- tribble(~ phylum, ~ bd_mass,
                   'Arthropoda', 0.2 + 1, # terrestrial + marine
                             # fish + livestock + humans + wild mammals + wild birds
                   'Chordata', 0.7 + 0.1 + 0.06 + 0.007 + 0.002,
                   'Annelida', 0.2, 
                   'Mollusca', 0.2, 
                   'Cnidaria', 0.1,
                   'Nematoda', 0.02) %>% 
           mutate(bd_total_biomass = sum(bd_mass), 
                  bd_prop_mass = bd_mass/bd_total_biomass) %>%
           left_join(mass_d, by='phylum')  %>%
           # not we need to reweight the original proportions as some phyla were
           # dropped
           mutate(prop_carbon_mass = prop_carbon_mass / sum(prop_carbon_mass))

ggplot(baron_d, aes(bd_prop_mass, prop_carbon_mass, color=phylum))  + 
  geom_point(size=5) + geom_abline(slope=1, intercept=0) + 
  xlab('Bar-On et al. 2018 proportion') + 
  ylab('My proportion')

@@

Arthropods are under sampled here, and chordates are way over sampled (at least
by biomass).

@@

ggplot(baron_d) + 
  geom_point(aes(phylum, total_carbon_mass), size=5)

ggplot(baron_d) + 
  geom_point(aes(phylum, bd_mass), size=5)

baron_d <- baron_d %>% mutate(prop_bd_biomass = total_carbon_mass / bd_mass)

baron_d %>%
  select(phylum, prop_bd_biomass)

>>>>

So, marine species are over represented in my dataset.

@@

A lot of the sources here are from the Bar-on et al. (2018) study.

Across arthropods and chordates, Baron-On et al use a wet weight to carbon
factor equal to 0.15. Description: "we convert wet weight values to carbon mass
assuming 70% water content and 50% carbon content out of dry weight."

For molluscs, cnidarians, nematodes, and annelids, they do something slightly
differently, often basing their conversion on the carbon of a single individual
(nematode, say). Here I forgo this and just use the conversion factor.

@@

@@

# https://bionumbers.hms.harvard.edu/bionumber.aspx?&id=107256 
# 18-23% of human body is carbon; his is close to the Bar-On et al. (2018) 
# estimate
prop_carbon <- 0.15 # from Bar-On et al.

# 1Gt = 1e12 kg
xd$carbon_mass <- (prop_carbon*(10^xd$pred_log10_body_mass) / 1e3 * # to kg
                   10^xd$log10_popsize) / 1e12

# take my data set and find the total carbon biomass of species, grouped by phyla
mass_d <- xd %>% 
  # filter so we don't include NA species in nspecies
  filter(is.finite(carbon_mass), is.finite(log10_popsize)) %>%
  group_by(phylum) %>% 
  summarize(total_carbon_mass=sum(carbon_mass, na.rm=TRUE),
            total_log10_popsize=log10(sum(10^log10_popsize, na.rm=TRUE)),
            nspecies = n()) %>%
  mutate(total = sum(total_carbon_mass), prop_carbon_mass = total_carbon_mass / total)

# chordates are incredibly over-represented in my data set by biomass
baron_d <- tribble(~ phylum, ~ bd_mass, 
                   'Arthropoda', 0.2 + 1, # terrestrial + marine
                             # fish + livestock + humans + wild mammals + wild birds
                   'Chordata', 0.7 + 0.1 + 0.06 + 0.007 + 0.002,
                   'Annelida', 0.2, 
                   'Mollusca', 0.2, 
                   'Cnidaria', 0.1,
                   'Nematoda', 0.02) %>% 
           mutate(bd_total_biomass = sum(bd_mass), 
                  bd_prop_mass = bd_mass/bd_total_biomass)

# merge my carbon masses in with Bar-On et al's
baron_mine <- baron_d %>%
           left_join(mass_d, by='phylum')  %>%
           # note, we need to reweight the original proportions as some phyla were
           # dropped because they are not in Bar-On's data set (e.g. Porifera, Echinoderms)
           mutate(prop_carbon_mass = prop_carbon_mass / sum(prop_carbon_mass))

@@

How are the proportions?

@@

ggplot(baron_mine, aes(bd_prop_mass, prop_carbon_mass, color=phylum))  + 
  geom_point(size=5) + geom_abline(slope=1, intercept=0) + 
  xlab('Bar-On et al. 2018 proportion') + 
  ylab('My proportion')

Arthropods are under sampled here, and chordates are way over sampled (at least
by biomass)!

Let's look at the biomass totals, in my small data set and Bar-On's:

ggplot(baron_mine) + 
  geom_point(aes(phylum, total_carbon_mass), size=5)

ggplot(baron_mine) + 
  geom_point(aes(phylum, bd_mass), size=5)

This shows the extent to which arthropods are under represented in my dataset
and chordates are overrepresented.

What fraction of the total biomasses in Bar-On are present in my dataset?

```{r}
baron_mine <- baron_mine %>% mutate(prop_bd_biomass = total_carbon_mass / bd_mass)

baron_mine %>% select(phylum, prop_bd_biomass)
```

4% of chordate biomass in my dataset seems high -- let's look at it.

First the top mammalian species by biomass:

xd %>% filter(class == 'Mammalia') %>% arrange(desc(carbon_mass)) %>% 
  select(species, carbon_mass) %>% 
  as.data.frame()

Baleen whales have a carbon mass of `0.0114e9 / 1e6 = 11.4` megatons! No way.

Actually this isn't wrong --- "In terms of their size and potential to store
carbon for years or decades, marine vertebrates are the only organisms in the
ocean comparable to large trees". This is from [this
paper](https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0012444)
which says ~10 megatons of carbon biomass is what we'd expect, pre-whaling.

mamd <- xd %>% filter(class == 'Mammalia') 
mamd$species  <- factor(mamd$species,
                        mamd$species[order(mamd$carbon_mass, decreasing=TRUE)], 
                        ordered=TRUE)
igno <- c("Balaenoptera bonaerensis", "Homo sapiens", "Eschrichtius robustus")
mamd %>% arrange(desc(carbon_mass)) %>% pull(species)

mamd %>% 
  filter(!(species %in% igno)) %>%
  ggplot(aes(species, carbon_mass))  + geom_point() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

xd %>% filter(species == "Aptenodytes patagonicus") %>% select(species, log10_popsize)

xd %>% filter(genus == "Equus") %>% select(species, log10_popsize, log10_range)

"Homo sapiens"
"Bos taurus"
"Ovis aries"
"Capra hircus"
